---
title: "Tutorial for making track plots to visualize finemapping results for genomic regions"
author: Kaixuan Luo
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

Track plot tutorial
=====================

**Required input data**:

* Genetic fine-mapping summary statistics.
* Gene annotations (exons, introns, UTRs, etc.).
* Functional annotation data, e.g.: 
ATAC-seq data, histone ChIP-seq peaks, PC-HiC loops, etc.

To make the trackplots, you will need to have the following packages installed:
`AnnotationDbi`, `org.Hs.eg.db`, `GenomicInteractions`, `Gviz`, `rtracklayer` from Bioconductor.

Load R packages
```{r load-packages, message=FALSE, warning=FALSE}
suppressMessages(library(data.table))
suppressMessages(library(tidyverse))
suppressMessages(library(GenomicRanges))
suppressMessages(library(AnnotationDbi)) # match gene ID to gene symbol
suppressMessages(library(org.Hs.eg.db)) # match gene ID to gene symbol
suppressMessages(library(GenomicInteractions)) # visualize HiC plots
suppressMessages(library(rtracklayer)) # loading bigwigs/bed files
suppressMessages(library(bigsnpr)) # loading genotype data from 1000Genomes for LD calculation
suppressMessages(library(Gviz)) # make track plots
library(mapgen)
source("../code/mapgen_trackplots.R")
```

```{r settings}
setwd("/project2/xinhe/kevinluo/gene-level-finemapping/trackplot_tutorial")
```

Load fine-mapping results.
```{r load-finemapping-res}
finemapstats <- readRDS(system.file("extdata", "aFib_Finemapped.tble.rds", package = "mapgen"))
finemapstats.gr <- process_finemapping_sumstats(finemapstats, 
                                                snp = 'snp', chr = 'chr', 
                                                pos = 'pos', pip = 'susie_pip', 
                                                pval = 'pval', zscore = 'zscore', 
                                                cs = 'CS', locus = 'locus',  
                                                pip.thresh = 0)
```

Load genomic annotations and gene information
```{r load-annotations}
genomic.annots <- readRDS(system.file("extdata", "genomic.annots.hg19.gr.rds", package = "mapgen"))
gene.annots <- genomic.annots$genes
```

Load Promoter-capture HiC (PCHi-C) data from iPSC derived cardiomyocytes (CMs).
```{r load-pcHiC}
pcHiC.gr <- readRDS(system.file("extdata", "pcHiC.CM.gr.rds", package = "mapgen"))
```

Load H3K27ac and DHS bed files
```{r load-peaks}
H3K27ac_peaks <- rtracklayer::import("data/H3K27ac_heart_concat.bed")
DHS_peaks <- rtracklayer::import("data/FetalHeart_E083-DNase_hg19_cleaned_narrowPeak.bed.gz")
```

Load ATAC data files. These data need to be in wig, bigWig/bw, bedGraph, or bam format. 
```{r load-counts-data}
atac_data_files <- c("data/Hg19_Cardiomyocyte-TileSize-500-normMethod-ReadsInTSS-ArchR.bw.bedGraph",
                     "data/Hg19_Endothelial-TileSize-500-normMethod-ReadsInTSS-ArchR.bw.bedGraph")

atac_data <- lapply(atac_data_files, function(x){rtracklayer::import(x)})
names(atac_data) <- c("Cardiomyocyte", "Endothelial")
```

Make a track plot to visualize a locus

Load gene mapping result
```{r load-gene-mapping-res}
gene.mapping.res <- readRDS("data/aFib_Finemapped_GeneMapped_ActivePromoter_07242021.gr.rds")
```

Get the genomic region for gene of interest
```{r gene-region}
gene.of.interest <- "FGF9"
region <- get_gene_region(gene.mapping.res, gene.of.interest, ext = 10000)
```

Make track plot
```{r track-plot, fig.width=12, fig.height=8, warning=FALSE}
pdf(paste0(gene.of.interest,'_trackplot.pdf'), width=12, height=8)
finemapping_annot_trackplot(finemapstats.gr, 
                            region, 
                            gene.annots, 
                            genome = "hg19", 
                            countsdata = atac_data, 
                            data_colors = c("red", "green"), 
                            data_ylim = c(0,0.8),
                            peaks = list("H3K27ac" = H3K27ac_peaks, "DHS" = DHS_peaks), 
                            HiC_loops = list("PC-HiC" = pcHiC.gr),
                            highlight_snps = "topSNP")
dev.off()
```

