---
title: "Finemapping using Mapgen with TORUS priors"
author: "Kaixuan Luo"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```

Here we show an example of performing enrichment analysis on AFib GWAS data using `mapgen` with `TORUS`. 

```{bash, eval=FALSE, include=FALSE}
cd /project2/xinhe/shared_data/mapgen/example_data/

mkdir -p example_data/GWAS/
cp /project2/gca/Heart_Atlas/reorganized_data/GWAS/summary_statistics/aFib/ebi-a-GCST006414_aFib.df.rds example_data/GWAS/ebi-a-GCST006414_aFib.df.rds

mkdir -p example_data/ATAC_seq/PeakCalls
cp ATAC_seq/PeakCalls/DA_MARKERS_FDRP_1_log2FC_1.rds example_data/ATAC_seq/PeakCalls/DA_MARKERS_FDRP_1_log2FC_1.rds

cp /project2/gca/aselewa/heart_atlas_project/GWAS/Torus/Torus_CellType_Enrichment_Results_Univariate_MORE.df.rds example_data/Torus/Torus_CellType_Enrichment_Results_Univariate_MORE.df.rds

cp /project2/xinhe/shared_data/aFib_gene_finemapping/aFib_Finemapped.tble.rds /project2/gca/Heart_Atlas/reorganized_data/example_data/finemapping/aFib_Finemapped.tble.rds

mkdir example_data/finemapping/Torus
```

```{r load-packages, eval=TRUE, message=FALSE}
library(mapgen)
library(tidyverse)
```

```{r setting}
data.dir <- '/project2/xinhe/shared_data/mapgen/example_data'
```

We use a reference genotype panel from European population (1KG). 
```{r load-bigSNP, eval=FALSE}
bigSNP <- bigsnpr::snp_attach(rdsfile = '/project2/xinhe/1kg/bigsnpr/EUR_variable_1kg.rds')
```

Load GWAS summary statistics of AFib
```{r load-GWAS, eval=FALSE}
gwas.sumstats <- readRDS(paste0(data.dir, '/GWAS/ebi-a-GCST006414_aFib.df.rds'))
gwas.sumstats <- gwas.sumstats %>% dplyr::rename(ss_index = og_index)
head(gwas.sumstats)
```

Prepare annotations for TORUS
```{r prepare-annots, eval=FALSE}
annotation_bed_files <- list.files(paste0(data.dir, '/finemapping/annotations_for_finemapping_hg19'), 
                                   pattern = '*.bed', full.names = T)
torus.files <- prepare_torus_input_files(gwas.sumstats, 
                                         annotation_bed_files, 
                                         torus_input_dir = paste0(data.dir, '/finemapping/torus_input'))
```

Run TORUS to estimate enrichment (joint annotations) and compute SNP-level prior

`run_torus()` with option = "est-prior" returns a list with: 
enrichment estimates (log odds ratio) and 95% confidence intervals of each annotation, 
and SNP-level priors using the enrichment estimates. 
```{r run-torus-est-prior, eval=FALSE}
torus.result <- run_torus(torus.files$torus_annot_file, 
                          torus.files$torus_zscore_file,
                          option = "est-prior",
                          torus_path = "torus") # set the path to 'torus' executable.
torus.enrich <- torus.result$enrich
torus.prior <- torus.result$snp_prior
saveRDS(torus.result, paste0(data.dir, '/finemapping/Torus_Enrichment_Results_Joint.rds'))
```

```{r}
torus.result <- readRDS(paste0(data.dir, '/finemapping/Torus_Enrichment_Results_Joint.rds'))
torus.enrich <- torus.result$enrich
torus.prior <- torus.result$snp_prior
```

Select GWAS significant loci with -log10(pval) < 5e-8
```{r select-gwas, eval=FALSE}
sig.loci <- gwas.sumstats %>% dplyr::filter(pval > -log10(5e-8)) %>% dplyr::pull(locus) %>% unique()
gwas.sumstats.sigloci <- gwas.sumstats[gwas.sumstats$locus %in% sig.loci, ]
cat(length(unique(gwas.sumstats.sigloci$locus)), "loci selected.\n")
```

Add Torus priors to GWAS summary statistics
```{r add-torus-prior, eval=FALSE}
sumstats.sigloci <- prepare_susie_data_with_torus_result(sumstats = gwas.sumstats.sigloci, 
                                                         torus_prior = torus.prior)
cat("Finemap",length(unique(sumstats.sigloci$locus)), "loci.\n")
saveRDS(sumstats.sigloci, paste0(data.dir, '/finemapping/sumstats.sigloci.rds'))
```

Run finemapping using SuSiE
```{r run-susie, eval=FALSE}
sumstats.sigloci <- readRDS(paste0(data.dir, '/finemapping/sumstats.sigloci.rds'))
cat("Finemapping",length(unique(sumstats.sigloci$locus)), "loci...\n")
# susie_finemap_L1 is a list of SuSiE results, one for each chunk/LD block. 
susie.res <- run_finemapping(sumstats.sigloci, bigSNP, priortype = 'torus', L = 1)

# add susie PIP information to GWAS summary stats
finemap.sumstats <- merge_susie_sumstats(susie.res, sumstats.sigloci)

saveRDS(finemap.sumstats, paste0(data.dir, '/finemapping/AF_finemapping_result_torusprior_122loci.rds'))
```

```{bash, include=FALSE, eval=FALSE}
sbatch --array=1-122 --time=1:0:0 --mem=15G /home/kaixuan/projects/analysis_pipelines/scripts/run_finemapping.sbatch
```


