---
title: "Enrichment analysis using mapgen with TORUS"
author: "Kaixuan Luo"
date: "2022-04-11"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

Here we show an example of performing enrichment analysis on AFib GWAS data using `mapgen` with `TORUS`. 

## Univariate enrichment analysis

Here we use scATAC-seq DA peaks for each cell type as a separate annotation (univariate).

This example is based on the [R script from Alan](https://github.com/xinhe-lab/heart_atlas/blob/master/R/torus_enrichment.R)

```{bash, eval=FALSE, include=FALSE}
cd /project2/gca/Heart_Atlas/reorganized_data/

mkdir -p example_data/GWAS/
cp /project2/gca/Heart_Atlas/reorganized_data/GWAS/summary_statistics/aFib/ebi-a-GCST006414_aFib.df.rds example_data/GWAS/ebi-a-GCST006414_aFib.df.rds

mkdir -p example_data/ATAC_seq/PeakCalls
cp ATAC_seq/PeakCalls/DA_MARKERS_FDRP_1_log2FC_1.rds example_data/ATAC_seq/PeakCalls/DA_MARKERS_FDRP_1_log2FC_1.rds

cp GWAS/Torus/Torus_CellType_Enrichment_Results_Univariate_MORE.df.rds example_data/Torus/Torus_CellType_Enrichment_Results_Univariate_MORE.df.rds
```

```{r load-packages, eval=TRUE, message=FALSE}
suppressMessages(library(liftOver))
suppressMessages(library(ComplexHeatmap))
library(mapgen)
```

```{r setting}
data.dir <- '/project2/gca/Heart_Atlas/reorganized_data/example_data'
```

Load GWAS summary statistics of AFib
```{r load-GWAS, eval=FALSE}
gwas.sumstats <- readRDS(paste0(data.dir, '/GWAS/ebi-a-GCST006414_aFib.df.rds'))
head(gwas.sumstats)
```

Prepare annotations for TORUS
```{r prepare-annots, eval=FALSE}
# load DA peaks (in hg38)
markers <- readRDS(paste0(data.dir, '/ATAC_seq/PeakCalls/DA_MARKERS_FDRP_1_log2FC_1.rds'))

# liftover peaks from hg38 to hg19
path <- system.file(package="liftOver", "extdata", "hg38ToHg19.over.chain")
ch <- import.chain(path)
markers.hg19 <- lapply(markers, function(x){unlist(liftOver(x, ch))})

system('mkdir -p Torus/bed_annotations_hg19')
# save to bed format
for(i in 1:length(markers.hg19)){
  seqlevelsStyle(markers.hg19[[i]]) <- "NCBI"
}

lapply(names(markers.hg19), function(x){
  rtracklayer::export(markers.hg19[[x]], 
                      format = 'bed', 
                      con = paste0(data.dir, '/Torus/bed_annotations_hg19/', x,'_narrowPeaks.bed'))})

annotations <- list.files(path = paste0(data.dir, '/Torus/bed_annotations_hg19'), pattern = '*.bed', full.names = T)

```

Run TORUS for each annotation separately
```{r run-torus, eval=FALSE}
enrich.res <- vector('list', length(annotations))
names(enrich.res) <- basename(annotations)

for(i in seq_along(annotations)){
  annot.name <- gsub('_narrowPeaks*', '', tools::file_path_sans_ext(basename(annotations[i])))

  # Make TORUS input data
  torus.files <- prepare_torus_input_files(gwas.sumstats, 
                                           annotations[i], 
                                           torus_input_dir = paste0(data.dir, '/Torus/input'))
  
  # Estimates enrichment using TORUS
  torus.result <- run_torus(torus.files$torus_annot_file,
                            torus.files$torus_zscore_file, 
                            option = "est", torus_path = "torus") 
  
  enrich.res[[i]] <- torus.result$enrich
}

# saveRDS(enrich.res, 'Torus/Torus_univariate_enrichment_result.rds')
```


Compare to pre-computed result
```{r, eval=FALSE}
enrich.alltraits.res <- readRDS(paste0(data.dir,'/Torus/Torus_CellType_Enrichment_Results_Univariate_MORE.df.rds'))
stopifnot(identical(enrich.res, enrich.alltraits.res$aFib))
```


### Plot enrichment for all traits

Load enrichment results
```{r load-extract-enrichment, eval=TRUE}

enrich.res <- readRDS(paste0(data.dir, '/Torus/Torus_CellType_Enrichment_Results_Univariate_MORE.df.rds'))
annotations <- list.files(path = paste0(data.dir, '/Torus/bed_annotations_hg19'), pattern = '*.bed', full.names = T)

pval_from_ci <- function(mean, upper, ci){
  nsamp <- length(mean)
  pval.res <- rep(0, nsamp)
  for(i in 1:nsamp){
    alph <- (1-ci)/2
    zval <- qnorm(p = 1-alph)
    se <- (upper[i]-mean[i])/zval
    
    pval.res[i] <- 1 - pnorm(q = mean[i] / se)   
  }
  return(pval.res)
}

res <- lapply(enrich.res, function(x){ Reduce(x = x, f = rbind)})
res <- lapply(res, function(x){x[x$term != "Intercept",]})
for(i in 1:length(res)){
  res[[i]]$pvalue <- pval_from_ci(mean = res[[i]]$estimate, upper = res[[i]]$high, ci = 0.95)
}
estimates <- as.data.frame(sapply(res, function(x){x["estimate"]}))
pvalues <- as.data.frame(sapply(res, function(x){x["pvalue"]}))
fdr <- matrix(p.adjust(unlist(pvalues), method = 'BH'), nrow = nrow(pvalues))

rnames <- basename(annotations)
names.order <- c("aFib", "PR_Interval","heart_rate","heart_failure",
                 "CAD","DiastolicBP","asthma","BMI","Height")
celltype_ideal_order <- c("Cardiomyocyte","Smooth Muscle","Pericyte","Endothelial","Fibroblast","Neuronal", "Lymphoid","Myeloid")
# celltype_ideal_order <- c("Cardiomyocyte","Pericyte","Endothelial","Fibroblast")

row.names(estimates) <- sub('_narrowPeaks.bed','',rnames)
colnames(estimates) <- names(enrich.res)
estimates <- estimates[celltype_ideal_order,names.order]
estimates <- t(estimates)

row.names(fdr) <- sub('_narrowPeaks.bed','',rnames)
colnames(fdr) <- names(enrich.res)
fdr <- fdr[celltype_ideal_order,names.order]
fdr <- t(fdr)

star.mat <- matrix('ns', nrow = nrow(fdr), ncol = ncol(fdr))
star.mat[fdr < 0.05] <- '*'
star.mat[fdr < 0.0001] <- '***'
rownames(star.mat) <- rownames(fdr)
colnames(star.mat) <- colnames(fdr)

mat.to.viz <- estimates/log(2)
mat.to.viz[mat.to.viz < 0] <- 0
```

Plot enrichment
```{r plot-heatmap, fig.width=5, fig.height=4, eval=TRUE}

lgd_list <- list()

col_fun <- c("lightblue", "orange", "firebrick")
names(col_fun) <- c("ns", '*', '***')

lgd_list[["fdr"]] <- Legend(title = "fdr (binned)",
                            labels = c("ns", '*', '***'),
                            legend_gp = gpar(fill = col_fun))

tic_vec <- c(0, 2, 4)
lgd_list[["log2_enrich"]] <- Legend(title = "log2_enrich",
                                    labels = tic_vec,
                                    # labels_gp = gpar(fontsize = 14),
                                    grid_height = unit(6, "mm"),
                                    grid_width = unit(6, "mm"),
                                    graphics = list(
                                      function(x, y, w, h) 
                                        grid.circle(x, y, 
                                                    r = (tic_vec[1]/10 + 0.2) * unit(2.5, "mm"),
                                                    gp = gpar(fill = "black")),
                                      function(x, y, w, h) 
                                        grid.circle(x, y, 
                                                    r = (tic_vec[2]/10 + 0.2) * unit(2.5, "mm"),
                                                    gp = gpar(fill = "black")),
                                      function(x, y, w, h) 
                                        grid.circle(x, y, 
                                                    r = (tic_vec[3]/10 + 0.2) * unit(2.5, "mm"),
                                                    gp = gpar(fill = "black"))
                                    ))

map1 <- Heatmap(star.mat,
                name = "Association Effect Size",
                col = col_fun,
                rect_gp = gpar(type = "none"),
                cell_fun = function(j, i, x, y, width, height, fill) {
                  grid.rect(x = x, y = y, width = width, height = height, 
                            gp = gpar(col = NA, fill = NA))
                  grid.circle(x = x, y = y,
                              r = (mat.to.viz[i, j]/10 + 0.2) * unit(2.5, "mm"),
                              gp = gpar(fill = col_fun[star.mat[i, j]], col = NA))
                },
                border_gp = gpar(col = "black"),
                row_title = "Trait",
                column_title = "Cell Type",
                cluster_rows = F, cluster_columns = F,
                show_heatmap_legend = F,
                row_names_gp = gpar(fontsize = 10.5),
                column_names_rot = 45,
                column_names_side = "top", 
                use_raster = T)

draw(map1, annotation_legend_list = lgd_list)
```

