---
title: "S-LDSC example analysis on GTEx eQTLs from Hormozdiari, F. et al. Nature Genetics 2018 paper"
author: "Kaixuan Luo"
date: "2020-07-07"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
## barplot for S-LDSC enrichment 
barplot_enrichment <- function(result_sLDSC, ylim = NULL, title = "", horizontal = FALSE){
  
  result_sLDSC$Enrichment_L <- result_sLDSC$Enrichment - 1.96*result_sLDSC$Enrichment_std_error
  result_sLDSC$Enrichment_H <- result_sLDSC$Enrichment + 1.96*result_sLDSC$Enrichment_std_error
  
  p <- ggplot(result_sLDSC, aes(x = Category, y = Enrichment))+
    geom_bar(position=position_dodge(), stat="identity", width = 0.5) +
    geom_errorbar(aes(ymin=Enrichment_L,
                      ymax=Enrichment_H),
                  width=.1,                    # Width of the error bars
                  position=position_dodge(.9)) + 
    ylab("Enrichment")+ xlab("") +
    ggtitle(title) +
    geom_hline(yintercept = 1,linetype="dotted", colour = "red")+
    theme_classic() +
    theme(axis.text.x = element_text(angle=30, hjust=1, size = 14))
  if(!is.null(ylim)){
    p <- p + coord_cartesian(ylim = ylim)
  }
  
  if(horizontal){
    p <- p + coord_flip()
  }
  print(p)
}

change_trait_names <- function(trait_namelist){
  trait_namelist <- gsub("PASS_","", trait_namelist)
  trait_namelist <- gsub("BMI1","BMI", trait_namelist)
  trait_namelist <- gsub("Rheumatoid_Arthritis","Rheumatoid Arthritis", trait_namelist)
  trait_namelist <- gsub("UKB_460K.blood_WHITE_COUNT","White Blood Cell Count", trait_namelist)
  trait_namelist <- gsub("UKB_460K.blood_PLATELET_COUNT","Platelet Count", trait_namelist)
}

```

## Run S-LDSC on data from Hormozdiari, F. et al. Nature Genetics 2018 paper
* Use the `GTEx_FE_META_TISSUE_GE_MaxCPP` annotation (MaxCPP annotation computed from fixed-effect meta-analysis of eQTLs from 44 GTEx tissues)
* Compute the `GTEx_FE_META_TISSUE_GE_MaxCPP` annotation conditional on `baselineLD_v1.1`
```{bash, eval=FALSE}

TRAITS=("PASS_BMI1" "PASS_Rheumatoid_Arthritis" "PASS_Schizophrenia" "UKB_460K.blood_WHITE_COUNT" "UKB_460K.blood_PLATELET_COUNT")

for trait in "${TRAITS[@]}"
do
  sbatch ~/projects/analysis_pipelines/code/sldsc_annot_GTEx_QTL_separate_example.sbatch ${trait} GTEx_FE_META_TISSUE_GE_MaxCPP
done

```

## Extract S-LDSC enrichment for the `GTEx_FE_META_TISSUE_GE_MaxCPP` annotation
* Note: the enrichment values of those traits should match those in the supplementary table 10 of Hormozdiari, F. et al. Nature Genetics 2018 paper.
```{r}
library(foreach)
library(doParallel)
registerDoParallel(cores = 6)

dir_results <- paste0("/project2/xinhe/kevinluo/ldsc/results/sLDSC_Hormozdiari_NG2018/LDSC_QTL/results_sLDSC/")

trait_name_list <- c("PASS_BMI1", "PASS_Rheumatoid_Arthritis", "PASS_Schizophrenia", "UKB_460K.blood_WHITE_COUNT", "UKB_460K.blood_PLATELET_COUNT")

prefix_annot <- "GTEx_FE_META_TISSUE_GE_MaxCPP"

result_sLDSC <- foreach(trait = trait_name_list, .combine = rbind)%dopar%{
  sldsc_results <- read.table(paste0(dir_results,"/", trait, "/baselineLDv1.1/", trait, "_", prefix_annot, "_baselineLDv1.1.results"), header = T, stringsAsFactors = F)
  est <- sldsc_results[sldsc_results$Category == "L2_1",]
  est$Category <- trait
  est
}

result_sLDSC$Category <- change_trait_names(result_sLDSC$Category)
DT::datatable(format(result_sLDSC, digits = 3), options = list(scrollX = TRUE, keys = TRUE, pageLength = 20),rownames = F)

```

### Plot S-LDSC enrichment for the `GTEx_FE_META_TISSUE_GE_MaxCPP` annotation
* Error bars represent 95% CI
```{r, fig.height=4, fig.width=4}
library(ggplot2)

barplot_enrichment(result_sLDSC)

```
